---
title: "Homework 6"
author: E. Brennan Bollman
date: December 9, 2020
output: github_document
---

```{r setup_and_data_visualization_preferences}
library(tidyverse)
library(patchwork)
library(stringr)
library(mgcv)
library(modelr)
library(purrr)
library(dplyr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

## Problem 1

#### Read and tidy data

```{r}
homicide_df = 
  read_csv("./data/homicide-data.csv") %>% 
  mutate(
    city_state = str_c(city, state, sep = ","),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest" ~ 0,
      disposition == "Closed by arrest" ~ 1
    )
  ) %>% 
  filter(city_state != "Tulsa,AL") %>% 
  filter(city_state != "Dallas,TX") %>% 
  filter(city_state != "Phoenix,AZ") %>% 
  filter(city_state != "Kansas City,MO") %>%
  filter(victim_race %in% c("White", "Black")) %>% 
  select(city_state, resolution, victim_age, victim_race, victim_sex)
```

#### Step 2: Attempt regression with one city

```{r}
baltimore_df = 
  homicide_df %>% 
  filter(city_state == "Baltimore,MD")

glm(resolution ~ victim_age + victim_race + victim_sex,
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```

#### Step 3: Iterate this logistic regression for each city

Will nest dataset, obtain df for each city, map glm over each city. Will tidy results as tibble.

```{r}
model_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x,
                         family = binomial())),
      results = map(models, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI"))
```

#### Step 4: Plot ORs for `victim_sex` for each city

```{r}
model_results_df %>%
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) + 
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1)) + 
  labs(
    title = "Odds Ratio of male vs female solved murders with 95% CIs",
    x = "City",
    y = "Odds ratio for male murders solved vs female murders (logistic regression)"
  )
```

Homicides in which the victim is male generally have OR estimate <1, in that male victims of homicide are less likely to be solved by arrest compared to female victims of homicide in most cities. By contrast, male-victim homicides are more likely to be solved by arrest than female-victim homicides in Albuqurque, Stockton CA, and Fresno (OR >1).


## Problem 2

#### Step 1: Read and tidy data

```{r}
birthweight_df = 
  read_csv("./data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  )
```

Note: in cleaning df, recoded categorical variables to factors, but intentionally left factors numeric rather than character because the numeric factor coding made sense. There are certainly problems with using "White" as a reference group; however, literature supports worse obstetric and neonatal outcomes for Black people and other people of color. Therefore, did not want these indicator variables to be coded alphabetically in model, and other approaches (such as based on frequency) did not make sense.

#### Step 2: Propose model 

To work through this problem in a stepwise manner, will trial building model based off prior medical learning (as an emergency physician) and quick literature searches. Must include `gaweeks`, `bhead` and `blength` since other baby size parameters likely related to birthweight.

Note: mother size incorporated using `ppbmi`. Did not include `ppwt` and `mheight` as too similar. Also, `pnumsga` felt to be too similar to `pnumlbw` so only one was used. 


```{r}
initial_model = lm(bwt ~ gaweeks + bhead + blength + babysex + delwt + fincome + frace + malform + momage + mrace + pnumlbw + ppbmi + smoken + wtgain, data = birthweight_df)

summary(initial_model)

broom::glance(initial_model)
```

Note: `pnumlbw` is NA in this model. When examining data, all observations are 0 for both `pnumlbw` and `pnumsga` so omitted these. 

Non-significant variables (removed): `fincome`, `frace`, `malform`, `momage`. `wtgain` is not highly significant, so will exclude also. `mrace` only significant for Black mothers versus White mothers (as expected), but will keep entire category.

```{r}
new_model = lm(bwt ~ gaweeks + bhead + blength + babysex + delwt + mrace + ppbmi + smoken, data = birthweight_df)

summary(new_model)

broom::tidy(new_model) %>% 
  select(-std.error, -statistic) %>% 
  knitr::kable()
```

Note: removing the above variables does not change R-squared. All values remain significant in new model.


#### Step 3: Plot Residuals

First look at distribution of residuals.

```{r}
birthweight_df %>% 
  add_residuals(new_model) %>% 
  ggplot(aes(x = resid)) + 
  geom_density()
```

Unfortunately, residuals have some right skew.

Next look at residuals versus predictors (parameter values).

```{r}
birthweight_df %>% 
  add_predictions(new_model) %>% 
  add_residuals(new_model) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point(alpha = 0.5)
```

There are definitely outliers in this distribution, but most points fit within a nice cluster, so will move forward with this model.

#### Step 4: Cross validation with other models

First, quickly look at other proposed models. 

```{r}
main_effects_model = lm(bwt ~ blength + gaweeks, data = birthweight_df)

broom::tidy(main_effects_model)

main_interactions_model = lm(bwt ~ babysex * bhead * blength, data = birthweight_df)

broom::tidy(main_interactions_model)
```

Next, build cross validation df. 

```{r}
cv_samp_df = 
  crossv_mc(birthweight_df, 100) %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))
```


```{r}
cross_validation_df = 
  cv_samp_df %>% 
  mutate(
    new_model = map(.x = train, ~lm(bwt ~ gaweeks + bhead + blength + babysex + delwt + mrace + ppbmi + smoken, data = .x)),
    main_effects_model = map(.x = train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    main_interactions_model = map(.x = train, ~lm(bwt ~ babysex * bhead * blength, data = .x))
  ) %>% 
  mutate(
    rmse_new_model = map2_dbl(.x = new_model, .y = test, ~rmse(model = .x, data = .y)),
    rmse_main_effects_model = map2_dbl(.x = main_effects_model, .y = test, ~rmse(model = .x, data = .y)),
    rmse_main_interactions_mosel = map2_dbl(.x = main_interactions_model, .y = test, ~rmse(model = .x, data = .y))
  )
```

#### Step 5: Compare models

Plot RMSE and compute averages

```{r}
cross_validation_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
```

```{r}
cross_validation_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  group_by(model) %>% 
  summarise(avg_rmse = mean(rmse)) %>% 
  knitr::kable()
```

The `new_model` created via rational, literature-based hypothesis of predictors of low birthweights performs best with the lowest rmse. Interesting, `main_interactions` also performs relatively well, but given that this model is not based off any rational hypothesis, would choose `new_model`. 


## Problem 3

Unfortunately, I am unable to get the RNOAA datasets. This has been a problem in class sessions before (though was previously able to use it on a HW once I think). Getting errors about unexpected ")" and "," essentially, even though the code is definitely correct (and often copied from the course website). Have tried what I could to troubleshoot.

I will review the solution to this problem for my learning.

I really appreciate this class and have learned a tremendous amount in a very short time! I feel confident in my ability to continue practicing and building these skills. Thanks for all your hard work teaching this class. The effort you put toward our learning was noticed :)